\input{macros_local}

In this chapter, we introduce and review results on stochastic games
with two players. On the one hand, they extend 2-player games with
random vertices; on the other hand, they extend Markov decision
processes with a second player. 

When equipped with a simple reachability objective, the objective of
\Eve is to maximize the probability of the set of plays that reach the
target, while here opponent has the opposite objective. Such games are
positionally determined: optimal positional pure strategies exist (see
Section~\ref{6-sec:values}). Perhaps surprisingly, stochastic games
with reachability objectives are central, in the sense that many more
complex objectives (discounted payoff, mean-payoff, parity) reduce to
them (see Section~\ref{6-sec:relations}). We therefore review several
algorithms to solve stochastic reachability games, based on the value
iteration or strategy improvement principles (see
Section~\ref{6-sec:algos}) that are also used for other classes of
games.



\section*{Notations}
\input{definitions}

\section{Positional determinacy of stochastic reachability games}
\label{6-sec:values}
\label{6-sec:determinacy}
% \begin{itemize}
% \item discussion qualitative vs quantitative: qualitative is a rather
%   small extension of non-stochastic games. We only need to encode
%   fairness into the winning condition, which is not too difficult
%   using a parity condition. Hence we will mostly focus on the
%   quantitative question.
% \item Speak of qualitative and quantitative determinacy
% \end{itemize}

\input{determinacy}

\section{Relations between all games}
\label{6-sec:relations}
\input{relations}

\section{Algorithmics}
\label{6-sec:algos}
\input{algos}

\section*{Bibliographic references}

\subsection{Pure memoryless determinacy of stochastic determinacy games}

\nat{A good reference for this proof is Kucera (chapter on stochastic
  games). Many papers reference Condon 92, but her proof less elegant,
  and from the beginning there is a restriction to MD
  strategies... The memoryless determinacy also holds for stochastic
  parity games (see CHJ-soda04). Check whether it is a byproduct of
  reductions.}

Originally papers assume (simple?) stopping games... for instance
Condon?