\input{macros_local}

This chapter considers concurrent games. The concurrent games we consider are extensions of the games considered in \cref{2-chap:regular} and \cref{4-chap:payoffs}, but where the choice of which edge to choose in a round is determined not by the choice of the owner of the vertex (indeed the vertices in concurrent games have no owners), but by the outcome of a matrix game corresponding to the vertex and played in that round. 
A matrix game is in turn a generalization of rock-paper-scissors, where each player picks an action simultaneously and then their pair of actions determines the outcome.

We will consider concurrent discounted, reachability and mean-payoff games and the definitions of the different objectives is as in the introduction. 
The chapter is divided into four sections:
\begin{enumerate}
\item The first section considers matrix games
\item The second section focuses on concurrent discounted games
\item The third section considers concurrent reachability games
\item The fourth section is about concurrent mean-payoff games
\end{enumerate}
As we go through the sections in this chapter, the complexity of the strategies and the computational complexity of solving the games rises: Indeed, since the games are generalizations of rock-paper-scissors, the strategies used requires randomness, but towards the end, no optimal or finite-memory $\epsilon$-optimal strategies exists in general and even the principle of sunken cost does not apply! 
Even with all this, the related questions about values are solvable in polynomial space and thus also in exponential time even in the last section.
The results we will focus on characterizes the complexity of the both the strategies as well as the computational complexity.
In each section we first give some positive result and some number of negative results. Each negative result also applies to the classes of games considered in the latter sections and each positive result applies to the classes considered in earlier sections (however, the positive results of latter sections will have worse complexity than the positive results from earlier sections).
As mentioned, the strategies for this chapter requires randomness and not too surprisingly, this implies that there is little difference between having stochastic or deterministic transition functions.

\section{Notations}
\label{7-sec:notations}
\input{notations}

\section{Matrix games}
\label{7-sec:matrix_games}
\input{matrix_games}

\section{Concurrent discounted games}
\label{7-sec:discounted}
\input{discounted}

\section{Concurrent reachability games}
\label{7-sec:reachability}
\input{reachability}

\section{Concurrent mean-payoff games}
\label{7-sec:mean_payoff}
\input{mean_payoff}

\section*{Bibilographic references}
\label{7-sec:references}
\input{references}








